National Pipeline System as a Solution to Future Water Shortage in the United States

### 2013-02-04 paper.tex
hey

\title{A National Pipeline System as a Solution to Future Water Shortage in the United States}

@@

In light of the 2012-2013 drought season, and the larger problem of growing demand for water across the nation, 
our department has conducted an
analysis of how best to satisfy the growing water needs of the United States.

@@

\section{The Case of Insufficient Funding}

Our model projects levels of supply and demand in 2025, and proceeds
to recommend infrastructure to meet that demand while minimizing cost.
This reflects the main precedent in major development of water supply
infrastructure in the United States, namely the \$24.6 billion Clean
Water Act \cite{key-15}, established with the aim of eliminating
all pollution of water supplies.

However, it is possible there will be insufficient funding for this
project. In this case, it will not be possible to construct all of
the recommended infrastructure. We approach this problem by creating
a priority listing, ordering the suggested developments in terms of
utility per dollar. To this end, we employ the following formula:

@@

\section{Avenues for Future Research}
A finer and more nuanced study of water usage and available data would provide a more complete model for water supplies and demands, as well as inform decisions about how to implement our proposals in reality. Here are some areas for future research:

\begin{itemize}
\item Groundwater level data is available in extreme detail from the US Geological Survey. Using this we can identify the areas of most severe exploitation as well as predict long-term aquifer lifespans. Areas with limited groundwater resources can then be prioritized in the building of the national pipelin network.
\item Research into currently available pipelines can be made if such data is collected; using these pipelines we predict that a large fraction of the network can simply be commandeered from existing infrastructure instead of built redundantly. This can be factored into the MST algorithm to give a minimal spanning tree based on pipelines already available. A problem with this approach is that available pipelines may not be of the condition or flow volume we require.
\item It is difficult with the granularity of the data available to actually model the optimal locations for specific de-salination plants and aqueducts to be built, although we can map out the general areas in which they are in need. The answer may perhaps be made on a case-by-case basis.
\item Reducing the total demand of the US is another approach that we did not address. New methods and technologies developed to this end should be used to inform our decisions about modeling the future demands of the US and decisions about pipeline-building.
\item Although we made the simplifying assumption that pipelines are the only largely available type of aqueduct, in reality in certain regions canals or other constructs may be more effective and less costly alternatives.
\end{itemize}

@@

The cost of the proposed infrastructure developments are significant, at a total of \$171 billion,
but must be placed in context. The cost of the 2012-2013 drought season
on the United States economy is estimated at between \$75 billion
and \$150 billion, and the potential total cost to the economy of
drought-related occurrences between 2013 and 2025 is much greater
even still.

>>>>

\title{National Pipeline System as a Solution to Future Water Shortage in the United States}

@@

In light of the 2012-2013 drought season, and the larger problem of growing demand for water across the nation, 
our department has conducted an
analysis of how best to satisfy the water needs of the United States.

@@

@@

\section{Potential Weaknesses}

Our proposed pipeline network can play a huge role in managing the distribution of freshwater over the United States. However, it is not a perfect solution to water management problems, and has some potential flaws:
\begin{itemize}
\item Before completion, there are holes in the network: Before every pipeline has been constructed, 
there will be some pairs of counties that are not connected by pipelines. In this case, the network is still 
useful, but is not quite as powerful as a complete network would be.
\item Large scale droughts can be an issue: if the entirety of the one region (say, the west coast) is in 
drought, while (say) the east coast has a water surplus, there may be bottlenecks in the middle of the country 
as large amounts of water are transferred. This could potentially be an issue, but in reality, this kind of situation 
is extremely rare, partially because the coasts of the United States do not tend to be where droughts occur.
\item Some nearby counties are only connected by long and convoluted paths: This is not as much of an issue as 
one might imagine, because generally droughts happen slowly, and hence it is acceptable for water to take a long time to 
reach its destination. However, if this is seen as a major issue, a small number of additional pipes can be added on a case-by-case basis.
\end{itemize}

However, none of these is a fatal flaw in the idea of constructing a pipeline network.

@@

The cost of the proposed infrastructure developments are significant, at a total of $171 billion,
but must be placed in context. The cost of the 2012-2013 drought season
on the United States economy is estimated at between \$75 billion
and \$150 billion, and the potential total cost to the economy of
drought-related occurrences between 2013 and 2025 is much greater
even still.
